\documentclass[aspectratio=169]{ISAE-Beamer}

%% Here comes your package calls!

\title[Short title]{A very very very very very very very, yes, very very very very very very very, yes, yes, very very very very, yes, yes, yes, very very very very, yes, yes, yes, yes, it's a very very very very very very very very very very long title}

\institute[ISAE, ONERA]
{\inst{1}ISAE-SUPAERO, Toulouse \and \inst{2}ONERA, Toulouse}

\author[JP, PP, PJ]{Jacques Pierre\inst{1} \and Pierre Paul\inst{2} \and Paul Jacques\inst{2}}

\date[THE Conf, 04/01/18]{\textit{The very very very best conference in the whole world, about something very serious}\\ April, the 1st, 2018}

\thanks{This work was granted by my holydays!!!\\ You're welcome!}

\begin{document}

\maketitle

\begin{frame}{Outline}

\tableofcontents

\end{frame}

\section{My first section}
\begin{frame}{This is a title}

\begin{block}{A simple block}
Something to say.
\end{block}

\begin{alertblock}{An important red block}
Some other thing to say.
\end{alertblock}

\begin{example}
An example block
\end{example}

\end{frame}

\begin{frame}

Frames without title are entitled as the current section.

\end{frame}

\subsection{A subsection}
\begin{frame}{This is a title}

With a defined title for the frame.

\end{frame}

\subsection{A second subsection}
\begin{frame}

with the current subsection's name as title.

\end{frame}

\section{Another section}
\begin{frame}

Just to see how it's working.
\vfill

We can also insert the logo of ISAE--SUPAERO with the command ``insertLogo'':

\begin{center}
\insertLogo
\end{center}
\vfill

It is vectorial and directly drawn using tikz in ``ISAE-Beamer.cls''.

Opening every folder in your computer to find this picture and losing a precious time? \begin{center}
\textbf{This is the past!}
\end{center}

\end{frame}

\begin{frame}{Thank you!}

\centering
\textbf{\Huge Questions?}

\end{frame}

\end{document}
