\documentclass[11pt]{letter}
 \usepackage[utf8]{inputenc} % un package
 \usepackage[T1]{fontenc} % un second package
 \usepackage[francais]{babel} % un troisième package
 %\usepackage{comment}
 \usepackage{textcomp}
 \usepackage{ifpdf}
\ifpdf
 \usepackage[pdftex]{graphicx}
 \else
 \usepackage[dvips]{graphicx}\fi
\pagestyle{empty}
\usepackage[top=0.5cm, bottom=0.5cm, left=1.5cm, right=1.5cm]{geometry}
 \setlength{\parindent}{0pt}
 \addtolength{\parskip}{6pt}
\renewcommand{\ttdefault}{pcr}
 \begin{document}
 %\sffamily
 %\hfill
 %
 \begin{flushleft}
 {\bfseries Andrea \textsc{BRUGNOLI}}\\[.35ex]
% \small\itshape
% 1 Avenue de Rangueil\\
% 31400, Toulouse\\[.35ex]
% 0033 7 50 39 47 27 \\
 Andrea.BRUGNOLI@isae.supaero.fr or andrea.brugnoli92@gmail.com
 \end{flushleft}
 %
 \begin{flushright}
 {\bfseries General Electric}\\[.35ex]
 \small\itshape
  Boston, Massachusetts, United States
  \end{flushright}
 %
 %\hfill
 %
 \begin{flushright}
 Toulouse, \today 
 \end{flushright}
 %
 \textbf{Subject}: Cover letter for Edison Engineering Development Program (EEDP)

 Dear Sir, Madame

General Electric reputation is daily confirmed by the growing number of successful projects. Its technical expertise is worldwide recognised and testified by the recognition that is one the most innovative (Fast Company, 2016) and admired (Fortune, 2016) companies world-wide. GE is known to be a thriving place for employees and leader. Focus on employees development and several leadership program allow GE's workforce to prosper and move forward in a rich and structured environment. GE stand over a solid foundation of integrity (for 11 straight years 2007–2017 it was honoured as one the most ethical companies), governance and sustainability. This latter is achieved in a wide sense at GE, incorporating the human, environmental and financial connotation.     Being able to participate to this successful and inspiring story and being surrounded by a stimulating, international and intensely engaged team are reasons of great enthusiasm and trepidation for me. This is why I am applying for the Edison Engineering program. 

 I specialised in mechanics and space engineering during my four years of study at Politecnico di Milano. I always engaged in my studies and thanks the achieved results I had the possibility to be involved in insightful experiences. During my second year of bachelor I could join an intensive formation in Labview, in partnership with National Instrument. The following year I worked on the kinematic and dynamic analysis of a forging manipulator. This project was carried out in partnership with Danieli Spa, an Italian multinational company. Our group was selected among twenty to give a final presentation in front the company engineers. 
 After my forth year I decided to engage in a new experience and to enrol myself at ISAE-Supaero. There I was challenged by a different pedagogic approach and destabilised by a different environment.  I learnt how to adapt myself and then found my path in the educational offer. Being inspired and amazed by how mathematics and automatics powers engineering and industry I decided to attend courses in applied mathematics  and advanced automatics. The former focus on deterministic optimisation, the latter on the dynamics of complex mechanical systems and model identification. My strong interest for the second one pushed me to enrol in a research master in automatics and image processing at university Paris Saclay Sud. I am now ending a six month long  internship at CNES, that provides me a critical attitude towards real engineering problem and future challenges.

I would love to apply my competences in a global, endlessly growing company like GE. The Edison Engineering Development Program could give me the possibility to express myself in a insightful contest, that will provide me a technical expertise, teach me efficient communication and  instill a sense of belonging to a great reality like GE. I am confident that I possess all is needed to work in a international and challenging environment.  I have no doubt that this program can enrich me in terms of technical capabilities and human values. In return I want to invest all my energies to benefit from this opportunity.

I stay at your disposal in case you might need contact me, at any time of your convenience.
Sincerely
 \begin{center}
 \large\textit{Andrea Brugnoli}
 \end{center}
 \end{document}

