\documentclass[11pt]{letter}
 \usepackage[utf8]{inputenc} % un package
 \usepackage[T1]{fontenc} % un second package
 \usepackage[francais]{babel} % un troisième package
 %\usepackage{comment}
 \usepackage{textcomp}
 \usepackage{ifpdf}
\ifpdf
 \usepackage[pdftex]{graphicx}
 \else
 \usepackage[dvips]{graphicx}\fi
\pagestyle{empty}
\usepackage[top=0.5cm, bottom=0.5cm, left=1.5cm, right=1.5cm]{geometry}
 \setlength{\parindent}{0pt}
 \addtolength{\parskip}{6pt}
\renewcommand{\ttdefault}{pcr}
 \begin{document}
 %\sffamily
 %\hfill
 %
 \begin{flushleft}
 {\bfseries Andrea \textsc{BRUGNOLI}}\\[.35ex]
% \small\itshape
% 1 Avenue de Rangueil\\
% 31400, Toulouse\\[.35ex]
% 0033 7 50 39 47 27 \\
 Andrea.BRUGNOLI@isae.supaero.fr or andrea.brugnoli92@gmail.com
 \end{flushleft}
 %
 \begin{flushright}
 {\bfseries ESA}\\[.35ex]
 \small\itshape
 Paris, Île-de-France, France
 \end{flushright}
 %
 %\hfill
 %
 \begin{flushright}
 Toulouse, \today 
 \end{flushright}
 %
 \textbf{Subject}: Cover letter for an internship position

 Dear Sir, Madame

ESA reputation is daily confirmed by the growing number of successful projects. Its technical expertise is worldwide recognised and testified by several collaborations with other international organisations, such as NASA or other important space agencies. The European space agency has always distinguished itself thanks to projects of huge scientific impact. Rosetta and Bepicolombo for example are among those and demonstrate ESA commitment towards its mission in the past and the present. Being able to contribute to projects of such importance and being surrounded by a stimulating, international and intensely engaged team are reasons of great enthusiasm and trepidation for me. This is why I am applying  for an internship at ESA. 

 I specialised in mechanics and space engineering during my four years of study at Politecnico di Milano. I always engaged in my studies and thanks the achieved results I had the possibility to be involved in insightful experiences. During my second year of bachelor I could join an intensive formation in Labview, in partnership with National Instrument. The following year I worked on the kinematic and dynamic analysis of a forging manipulator. This project was carried out in partnership with Danieli Spa, an Italian multinational company. Our group was selected among twenty to give a final presentation in front the company engineers. 
 %I especially enjoyed working on an interplanetary transfer project. In  challenging space missions many different disciplines have to be integrated and this is why I find it an intriguing subject. Later on I had the occasion to work on multi-body dynamics in Simulink environment to study appendices flexibilities during satellites manoeuvers.  All those projects provided me a pragmatic formation combined with a theoretical background. 
 After my forth year I decided to engage in a new experience and to enrol myself at ISAE-Supaero. There I was challenged by a different pedagogic approach and destabilised by a different environment.  I learnt how to adapt myself and then found my path in the educational offer. I have always been amazed by how math can unveil the mysteries of nature. At the same time engineering is the medium by which we can obverse math potentialities at work. In particular automatics keeps changing and improving our world  and way of living. These are reasons why I am now attending courses in applied mathematics  and advanced automatics. The former focus on deterministic optimisation, the latter on the dynamics of complex mechanical systems and model identification. My strong interest for the second one pushed me to enrol in a research master in automatics and image processing at university Paris Saclay Sud.

I would love to apply my competences in the space field. Space applications contribute to worldwide cooperation.  They had, have and will always have a key role in developing crucial technologies. Space research is the highest frontier of human progress and will allow mankind to discover before inconceivable mission scenarios. ESA could give me the possibility to express myself in a insightful contest. I am confident that I possess all is needed to work in a international environment.  I have no doubt that an internship at ESA can enrich me in terms of technical capabilities and human values. In return I want to invest all my energies to benefit from this opportunity.

I stay at your disposal in case you might need contact me, at any time of your convenience.
Sincerely
 \begin{center}
 \large\textit{Andrea Brugnoli}
 \end{center}
 \end{document}

